---
title: "Download data from the County Business Pattern (CBP)"
author: "Erik Loualiche"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CBP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

We download data from the CBP. Data is available for the period from 1986 to 2013 at an annual frequency.
The most detailed cut of the data incorporates size, country and industry (4 digits) employment and businesses. 

For example if we would like the raw data:
```{r}
dt_raw <- download_all_cbp(year_start, year_end)
dt_raw
```

For something a bit tidier we gather the data around year, geography, industry and size:
```{r}
dt_tidy <- tidy_cbp(dt_raw, year_start, year_end)
dt_tidy
```

### Some more information about data layout

----------------------------------------------------------------
| Variable Name             | Geographic/Size LevelCode Format |
|---------------------------|----------------------------------|
| Size                      |                                  |
|                           |                                  |
| `N1_4`                    | 1-4 Employee Size Class          |
| `N5_9`                    | 5-9 Employee Size Class          |
| `N10_19`                  | 10-19 Employee Size Class        |
| `N20_49`                  | 20-49 Employee Size Class        |
| `N50_99`                  | 50-99 Employee Size Class        |
| `N100_249`                | 100-249 Employee Size Class      |
| `N250_499`                | 250-499 Employee Size Class      |
| `N500_999`                | 500-999 Employee Size Class      |
| `N1000`                   | 1000 or More Employee Size Class |
| `N1000_1`                 | 1000-1499 Employee Size Class    |
| `N1000_2`                 | 1500-2499 Employee Size Class    |
| `N1000_3`                 | 2500-4999 Employee Size Class    |
| `N1000_4`                 | 5000 or More Employee Size Class |
|                           |                                  |
| 
