---
title: "Download data from the Quarterly Census of Employment and Wages (QCEW)"
author: "Erik Loualiche"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
#output:
#  md_document:
#    variant: markdown_github
#    toc: true
#    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Download data from the Quarterly Census of Employment and Wages (QCEW)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r global_options, echo = FALSE, include = FALSE}
options(width = 999)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      cache = FALSE, tidy = FALSE, size = "small")
```

In this vignette we provide examples of how to use entrydatar to download precise files from the QCEW.     
In the second part we give rudimentary documentations of the feature that are of principal interest to work with the data.
For an overview of the QCEW, head over to the [BLS website](https://www.bls.gov/cew/cewover.htm)

The vignette is organized as:

1. Downloading data from the QCEW
2. Documentation
   + Table of data cuts: <sup>[here](#tablecut)</sup>
   + General documentation: <sup>[here](#generaldoc)</sup>




# Downloading data from the QCEW

To get read to download data, we load some libraries that the package might have forgotten to call (the package works with all tables in a `data.table` format)
```{r, results='hide', warning = F, error = F, message = F}
library(dplyr); library(data.table); 
library(entrydatar); 
```

## General download

For example if we are interested in downloading aggregate level data we use the cut `10`. Please see documentation below for a definition of what is availabe and the mapping of the different cuts of QCEW.
```{r, results='hide', warning = F, error = F, message = F, eval = F}
dt_agg <- get_files_cut(       
    data_cut = 10,     
    year_start = 1990, year_end =1993,    
    path_data = "~/Downloads/", write = F)
```
```{r, eval = F}
as_tibble(dt_agg)
# A tibble: 16 x 42
   area_fips own_code indus… agglv… size_…  year   qtr discl… qtrly… month… month… month… total_… taxa… qtrl… avg_… lq_d… lq_q… lq_m… lq_m… lq_m… lq_t… lq_t… lq_q… lq_a… oty_… oty_q… oty_… oty_mo… oty_m… oty_mo… oty_m… oty_mo… oty_m… oty_to… oty_t… oty_tax… oty_… oty_qt… oty_… oty_a… oty_…
   <chr>        <dbl>  <dbl>  <dbl>  <dbl> <dbl> <dbl> <chr>   <dbl>  <dbl>  <dbl>  <dbl>   <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>  <dbl> <dbl>   <dbl>  <dbl>   <dbl>  <dbl>   <dbl>  <dbl>   <dbl>  <dbl>    <dbl> <dbl>   <dbl> <dbl>  <dbl> <dbl>
 1 US000            0   10.0   10.0      0  1990  1.00 ""     6.02e⁶ 1.06e⁸ 1.07e⁸ 1.08e⁸ 6.23e¹¹     0     0   448    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA 127332  2.20  2.09e⁶  2.00   2.12e⁶  2.00   2.03e⁶  1.90  3.91e¹⁰  6.70  -3.57e¹¹  -100 -7.56e⁹  -100  20.0   4.70
 2 US000            0   10.0   10.0      0  1990  2.00 ""     6.04e⁶ 1.08e⁸ 1.09e⁸ 1.10e⁸ 6.29e¹¹     0     0   443    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA 154120  2.60  1.87e⁶  1.80   2.12e⁶  2.00   1.82e⁶  1.70  3.67e¹⁰  6.20  -1.82e¹¹  -100 -4.01e⁹  -100  19.0   4.50
 3 US000            0   10.0   10.0      0  1990  3.00 ""     6.09e⁶ 1.08e⁸ 1.09e⁸ 1.10e⁸ 6.24e¹¹     0     0   441    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA 164809  2.80  1.68e⁶  1.60   1.79e⁶  1.70   1.34e⁶  1.20  3.37e¹⁰  5.70  -1.10e¹¹  -100 -2.47e⁹  -100  18.0   4.30
 4 US000            0   10.0   10.0      0  1990  4.00 ""     6.12e⁶ 1.09e⁸ 1.09e⁸ 1.09e⁸ 6.88e¹¹     0     0   483    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA 149102  2.50  1.01e⁶  0.900  6.69e⁵  0.600  2.75e⁵  0.300 3.90e¹⁰  6.00  -8.46e¹⁰  -100 -1.85e⁹  -100  25.0   5.50
 5 US000            0   10.0   10.0      0  1991  1.00 ""     6.30e⁶ 1.05e⁸ 1.05e⁸ 1.06e⁸ 6.37e¹¹     0     0   465    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA 284012  4.70 -1.03e⁶ -1.00  -1.51e⁶ -1.40  -1.77e⁶ -1.60  1.40e¹⁰  2.20   0           0  0          0  17.0   3.80
 6 US000            0   10.0   10.0      0  1991  2.00 ""     6.36e⁶ 1.06e⁸ 1.07e⁸ 1.08e⁸ 6.45e¹¹     0     0   463    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA 317123  5.30 -1.84e⁶ -1.70  -2.05e⁶ -1.90  -2.24e⁶ -2.00  1.64e¹⁰  2.60   0           0  0          0  20.0   4.50
 7 US000            0   10.0   10.0      0  1991  3.00 ""     6.39e⁶ 1.06e⁸ 1.07e⁸ 1.08e⁸ 6.41e¹¹     0     0   462    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA 303568  5.00 -2.13e⁶ -2.00  -1.93e⁶ -1.80  -2.03e⁶ -1.90  1.76e¹⁰  2.80   0           0  0          0  21.0   4.80
 8 US000            0   10.0   10.0      0  1991  4.00 ""     6.43e⁶ 1.08e⁸ 1.08e⁸ 1.08e⁸ 7.05e¹¹     0     0   503    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA 317664  5.20 -1.70e⁶ -1.60  -1.61e⁶ -1.50  -1.52e⁶ -1.40  1.72e¹⁰  2.50   0           0  0          0  20.0   4.10
 9 US000            0   10.0   10.0      0  1992  1.00 ""     6.48e⁶ 1.05e⁸ 1.05e⁸ 1.06e⁸ 6.64e¹¹     0     0   486    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA 176781  2.80 -5.04e⁵ -0.500 -1.48e⁵ -0.100 -1.06e⁵ -0.100 2.72e¹⁰  4.30   0           0  0          0  21.0   4.50
10 US000            0   10.0   10.0      0  1992  2.00 ""     6.50e⁶ 1.07e⁸ 1.08e⁸ 1.09e⁸ 6.72e¹¹     0     0   479    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA 140624  2.20  5.39e⁵  0.500  6.40e⁵  0.600  6.08e⁵  0.600 2.69e¹⁰  4.20   0           0  0          0  16.0   3.50
11 US000            0   10.0   10.0      0  1992  3.00 ""     6.52e⁶ 1.07e⁸ 1.07e⁸ 1.08e⁸ 6.72e¹¹     0     0   482    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA 131248  2.10  7.06e⁵  0.700  2.48e⁵  0.200  6.40e⁵  0.600 3.06e¹⁰  4.80   0           0  0          0  20.0   4.30
12 US000            0   10.0   10.0      0  1992  4.00 ""     6.57e⁶ 1.09e⁸ 1.09e⁸ 1.09e⁸ 7.79e¹¹     0     0   550    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA 134740  2.10  1.09e⁶  1.00   1.05e⁶  1.00   1.22e⁶  1.10  7.36e¹⁰ 10.4    0           0  0          0  47.0   9.30
13 US000            0   10.0   10.0      0  1993  1.00 ""     6.62e⁶ 1.06e⁸ 1.07e⁸ 1.07e⁸ 6.67e¹¹     0     0   480    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA 138525  2.10  1.43e⁶  1.40   1.70e⁶  1.60   1.69e⁶  1.60  2.54e ⁹  0.400  0           0  0          0 - 6.00 -1.20
14 US000            0   10.0   10.0      0  1993  2.00 ""     6.65e⁶ 1.09e⁸ 1.10e⁸ 1.11e⁸ 7.04e¹¹     0     0   494    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA 155234  2.40  1.85e⁶  1.70   1.75e⁶  1.60   1.98e⁶  1.80  3.19e¹⁰  4.70   0           0  0          0  15.0   3.10
15 US000            0   10.0   10.0      0  1993  3.00 ""     6.70e⁶ 1.09e⁸ 1.09e⁸ 1.11e⁸ 7.11e¹¹     0     0   499    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA 176159  2.70  2.15e⁶  2.00   2.31e⁶  2.20   2.35e⁶  2.20  3.86e¹⁰  5.70   0           0  0          0  17.0   3.50
16 US000            0   10.0   10.0      0  1993  4.00 ""     6.75e⁶ 1.11e⁸ 1.11e⁸ 1.12e⁸ 8.06e¹¹     0     0   556    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA 176528  2.70  2.42e⁶  2.20   2.54e⁶  2.30   2.80e⁶  2.60  2.76e¹⁰  3.50   0           0  0          0   6.00  1.10
```

Note that the dataset can be large. Downloading most of the industry 3 and 4 digits cuts across counties and MSAs ends up with 57mn rows. Saving it in `.rds` formats takes around 2gb of memory.

On the other hand for something more precise, say nationally by size cuts at the 6 digits industry level we would call the cut `28`:
```{r, results='hide', warning = F, error = F, message = F, eval = F}
dt_naics <- get_files_cut(        
  data_cut = 28,       
  industry = "naics", 
  year_start = 1990, year_end = 2015,       
  path_data = "~/Downloads/", write = F)
```
```{r, eval = F}
as_tibble(dt_naics)
# A tibble: 208,667 x 47
   area… own_… indus… aggl… size…  year   qtr disc… area… own_… indu… aggl… size… qtrly… month… month… month… total… taxa… qtrl… avg_… lq_d… lq_q… lq_m… lq_m… lq_m… lq_t… lq_t… lq_q… lq_a… oty_… oty_… oty_… oty_… oty_… oty_… oty_… oty_… oty_… oty_… oty_… oty_… oty_… oty_… oty_… oty_… oty_…
   <chr> <dbl>  <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>
 1 US000  5.00 111110  28.0  1.00  1990  1.00    NA    NA    NA    NA    NA    NA 2.16e²  354    353    371   1.57e⁶     0     0   337    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA     0     0     0     0     0     0     0     0     0     0     0     0     0     0     0     0
 2 US000  5.00 111110  28.0  2.00  1990  1.00    NA    NA    NA    NA    NA    NA 9.80e¹  547    577    640   1.95e⁶     0     0   255    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA     0     0     0     0     0     0     0     0     0     0     0     0     0     0     0     0
 3 US000  5.00 111110  28.0  3.00  1990  1.00    NA    NA    NA    NA    NA    NA 4.00e¹  439    450    519   1.32e⁶     0     0   216    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA     0     0     0     0     0     0     0     0     0     0     0     0     0     0     0     0
 4 US000  5.00 111110  28.0  4.00  1990  1.00    NA    NA    NA    NA    NA    NA 1.40e¹  336    336    382   1.01e⁶     0     0   220    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA     0     0     0     0     0     0     0     0     0     0     0     0     0     0     0     0
 5 US000  5.00 111120  28.0  1.00  1990  1.00    NA    NA    NA    NA    NA    NA 2.60e¹   39.0   33.0   39.0 1.30e⁵     0     0   270    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA     0     0     0     0     0     0     0     0     0     0     0     0     0     0     0     0
 6 US000  5.00 111120  28.0  2.00  1990  1.00    NA    NA    NA    NA    NA    NA 5.00e⁰   21.0   23.0   28.0 6.87e⁴     0     0   220    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA     0     0     0     0     0     0     0     0     0     0     0     0     0     0     0     0
 7 US000  5.00 111130  28.0  1.00  1990  1.00    NA    NA    NA    NA    NA    NA 8.20e¹   76.0   73.0   85.0 3.28e⁵     0     0   324    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA     0     0     0     0     0     0     0     0     0     0     0     0     0     0     0     0
 8 US000  5.00 111140  28.0  1.00  1990  1.00    NA    NA    NA    NA    NA    NA 1.92e³ 1369   1314   1575   5.09e⁶     0     0   276    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA     0     0     0     0     0     0     0     0     0     0     0     0     0     0     0     0
 9 US000  5.00 111140  28.0  2.00  1990  1.00    NA    NA    NA    NA    NA    NA 8.50e¹  374    404    526   1.34e⁶     0     0   237    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA     0     0     0     0     0     0     0     0     0     0     0     0     0     0     0     0
10 US000  5.00 111140  28.0  4.00  1990  1.00    NA    NA    NA    NA    NA    NA 8.00e⁰  193    211    224   7.45e⁵     0     0   274    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA     0     0     0     0     0     0     0     0     0     0     0     0     0     0     0     0
# ... with 208,657 more rows
```


## Faster download with annual frequencies

The BLS provides quarterly as well as annual averages, which are lighter. There is now an option to download directly the annual average using a frequency option. The default is quarterly, so everything should be backward compatible.
Let's see first with the quarterly data (*warning* it is slow):
```{r, results='hide', warning = F, error = F, message = F, eval = F}
dt_naics <- get_files_cut(     
  data_cut = 16,     
  industry = "naics",  frequency = "quarter",  
  year_start = 2000, year_end = 2005,    
  path_data = "~/Downloads/", write = F)
```
```{r, eval = F}
as_tibble(dt_naics)
# A tibble: 18,028 x 42
   area_fips own_code indus… agglv… size_…  year   qtr discl… qtrly_… month… month… month… total… taxa… qtrl… avg_… lq_d… lq_q… lq_m… lq_m… lq_m… lq_t… lq_t… lq_q… lq_a… oty_… oty_… oty_q… oty_mo… oty_mo… oty_mo… oty_m… oty_mo… oty_m… oty_to… oty_to… oty_… oty_… oty_… oty_… oty_av… oty_av…
   <chr>        <dbl>  <dbl>  <dbl>  <dbl> <dbl> <dbl> <chr>    <dbl>  <dbl>  <dbl>  <dbl>  <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>  <dbl>   <dbl>   <dbl>   <dbl>  <dbl>   <dbl>  <dbl>   <dbl>   <dbl> <dbl> <dbl> <dbl> <dbl>   <dbl>   <dbl>
 1 US000         1.00   1125   16.0      0  2000  1.00 ""        2.00 7.00e⁰ 7.00e⁰ 7.00e⁰ 5.24e⁴     0     0   576    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA  0      0    -  1.00 -12.5   -  1.00 -12.5  -  1.00 -12.5  -1.90e³ - 3.50      0     0     0     0   54.0   10.3  
 2 US000         1.00   1125   16.0      0  2000  2.00 ""        2.00 7.00e⁰ 7.00e⁰ 8.00e⁰ 5.59e⁴     0     0   587    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA  0      0    -  1.00 -12.5      0      0       1.00  14.3   1.26e⁴  29.0       0     0     0     0  132     29.0  
 3 US000         1.00   1125   16.0      0  2000  3.00 ""        2.00 8.00e⁰ 8.00e⁰ 8.00e⁰ 5.99e⁴     0     0   576    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA  0      0       1.00  14.3      1.00  14.3     1.00  14.3   9.74e³  19.4       0     0     0     0   25.0    4.50 
 4 US000         1.00   1125   16.0      0  2000  4.00 ""        3.00 8.00e⁰ 8.00e⁰ 8.00e⁰ 6.42e⁴     0     0   617    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA  1.00  50.0     1.00  14.3      1.00  14.3     1.00  14.3   1.34e⁴  26.4       0     0     0     0   59.0   10.6  
 5 US000         1.00   1131   16.0      0  2000  1.00 ""       78.0  4.09e³ 3.92e³ 3.93e³ 3.66e⁷     0     0   706    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA -1.00 - 1.30 -237    - 5.50  -217    - 5.20 -282    - 6.70 -1.22e⁵ - 0.300     0     0     0     0   38.0    5.70 
 6 US000         1.00   1131   16.0      0  2000  2.00 ""       78.0  4.19e³ 4.57e³ 5.26e³ 4.77e⁷     0     0   786    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA  1.00   1.30 -315    - 7.00  -255    - 5.30  250      5.00  7.70e⁶  19.2       0     0     0     0  142     22.0  
 7 US000         1.00   1131   16.0      0  2000  3.00 ""       78.0  5.85e³ 5.67e³ 5.47e³ 5.86e⁷     0     0   796    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA  0      0    - 97.0  - 1.60  -167    - 2.90 -110    - 2.00 -1.43e⁶ - 2.40      0     0     0     0 -  1.00 - 0.100
 8 US000         1.00   1131   16.0      0  2000  4.00 ""       78.0  4.84e³ 4.49e³ 4.43e³ 5.09e⁷     0     0   853    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA  0      0    -502    - 9.40  -146    - 3.10  205      4.80 -1.07e⁵ - 0.200     0     0     0     0   25.0    3.00 
 9 US000         1.00   2211   16.0      0  2000  1.00 ""      140    1.23e⁴ 1.23e⁴ 1.23e⁴ 1.91e⁸     0     0  1193    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA -5.00 - 3.40 -201    - 1.60  -157    - 1.30 -125    - 1.00  1.54e⁷   8.80      0     0     0     0  110     10.2  
10 US000         1.00   2211   16.0      0  2000  2.00 ""      141    1.24e⁴ 1.24e⁴ 1.24e⁴ 1.83e⁸     0     0  1138    NA  1.00  1.00  1.00  1.00  1.00     0     0  1.00    NA -2.00 - 1.40   49.0    0.400    5.00   0       6.00   0     2.54e⁷  16.1       0     0     0     0  157     16.0  
# ... with 18,018 more rows
```

Now let's download the data only at annual frequency (similar to the size cuts from earlier):
```{r, results='hide', warning = F, error = F, message = F, eval = F}
dt_naics_year <- get_files_cut(     
  data_cut = 16,     
  industry = "naics",  frequency = "year",  
  year_start = 2000, year_end = 2005,    
  path_data = "~/Downloads/", write = F)
```
```{r, eval = F}
as_tibble(dt_naics_year)
# A tibble: 4,507 x 38
   area_fips own_code industry_code agglvl_code size_code  year   qtr disclosure_code annual… annual… total_… taxab… annua… annua… avg_a… lq_di… lq_a… lq_a… lq_t… lq_t… lq_a… lq_a… lq_a… oty_… oty_… oty_a… oty_an… oty_ann… oty_to… oty_tot… oty_… oty_… oty_… oty_… oty_a… oty_a… oty_… oty_a…
   <chr>        <dbl>         <dbl>       <dbl>     <dbl> <dbl> <dbl> <chr>             <dbl>   <dbl>   <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>  <dbl>   <dbl>    <dbl>   <dbl>    <dbl> <dbl> <dbl> <dbl> <dbl>  <dbl>  <dbl> <dbl>  <dbl>
 1 US000         1.00          1125        16.0         0  2000    NA ""                 2.00  8.00e⁰  2.32e⁵      0      0    589  30648     NA  1.00  1.00  1.00     0     0  1.00  1.00    NA  0      0     1.00e⁰   14.3    3.38e⁴   17.0       0     0     0     0   68.0  13.1   3567  13.2 
 2 US000         1.00          1131        16.0         0  2000    NA ""                78.0   4.73e³  1.94e⁸      0      0    788  40994     NA  1.00  1.00  1.00     0     0  1.00  1.00    NA  0      0    -1.56e² -  3.20   6.04e⁶    3.20      0     0     0     0   49.0   6.60  2547   6.60
 3 US000         1.00          2211        16.0         0  2000    NA ""               142     1.24e⁴  7.90e⁸      0      0   1222  63557     NA  1.00  1.00  1.00     0     0  1.00  1.00    NA -2.00 - 1.40 -5.10e¹ -  0.400  8.53e⁷   12.1       0     0     0     0  136    12.5   7096  12.6 
 4 US000         1.00          2213        16.0         0  2000    NA ""                10.0   9.37e²  2.67e⁷      0      0    547  28463     NA  1.00  1.00  1.00     0     0  1.00  1.00    NA  1.00  11.1   1.14e²   13.9   -4.72e⁶ - 15.0       0     0     0     0 -186   -25.4  -9643 -25.3 
 5 US000         1.00          2362        16.0         0  2000    NA ""                 1.00  1.00e⁰  6.73e⁴      0      0   1294  67270     NA  1.00  1.00  1.00     0     0  1.00  1.00    NA  0      0     1.00e⁰  100      4.32e⁴  180         0     0     0     0  184    16.6   9555  16.6 
 6 US000         1.00          3231        16.0         0  2000    NA ""                 8.00  4.86e³  2.69e⁸      0      0   1063  55287     NA  1.00  1.00  1.00     0     0  1.00  1.00    NA  0      0    -6.10e¹ -  1.20   8.19e⁵    0.300     0     0     0     0   16.0   1.50   852   1.60
 7 US000         1.00          3259        16.0         0  2000    NA ""                 3.00  3.10e²  1.67e⁷      0      0   1037  53945     NA  1.00  1.00  1.00     0     0  1.00  1.00    NA  0      0    -2.00e⁰ -  0.600  1.30e⁶    8.40      0     0     0     0   84.0   8.80  4396   8.90
 8 US000         1.00          3321        16.0         0  2000    NA ""                 3.00  4.75e²  3.06e⁷      0      0   1241  64540     NA  1.00  1.00  1.00     0     0  1.00  1.00    NA  1.00  50.0   1.26e²   36.1    9.43e⁶   44.5       0     0     0     0   72.0   6.20  3748   6.20
 9 US000         1.00          3329        16.0         0  2000    NA ""                10.0   1.23e⁴  6.24e⁸      0      0    976  50746     NA  1.00  1.00  1.00     0     0  1.00  1.00    NA  0      0    -1.32e³ -  9.70   5.99e⁶    1.00      0     0     0     0  104    11.9   5379  11.9 
10 US000         1.00          3364        16.0         0  2000    NA ""                 1.00  2.60e³  6.86e⁷      0      0    508  26390     NA  1.00  1.00  1.00     0     0  1.00  1.00    NA  0      0    -4.50e¹ -  1.70  -1.19e⁷ - 14.8       0     0     0     0 - 77.0 -13.2  -4048 -13.3 
# ... with 4,497 more rows
```


## Tidy the data

The data is still raw from the BLS download. We can clean the data in two ways: a monthly dataset with employment by categories and a quarterly dataset with only number of firms. Finally we allow an option for aggregating both types. Note this function works only for the quarterly version of the data.
```{r, results='hide', warning = F, error = F, message = F, eval = F}
dt_naics_tidy <- tidy_qcew(dt_naics, frequency = "month")
```
```{r, eval = F}
as_tibble(dt_naics_tidy)
# A tibble: 54,084 x 9
    year quarter industry_code area_fips own_code size_code agglvl_code month emplvl
   <dbl>   <dbl>         <dbl> <chr>        <dbl>     <dbl>       <dbl> <dbl>  <dbl>
 1  2000    1.00          1111 US000         5.00         0        16.0  1.00  21243
 2  2000    1.00          1111 US000         5.00         0        16.0  2.00  21759
 3  2000    1.00          1111 US000         5.00         0        16.0  3.00  23600
 4  2000    2.00          1111 US000         5.00         0        16.0  4.00  26142
 5  2000    2.00          1111 US000         5.00         0        16.0  5.00  27287
 6  2000    2.00          1111 US000         5.00         0        16.0  6.00  29355
 7  2000    3.00          1111 US000         5.00         0        16.0  7.00  39098
 8  2000    3.00          1111 US000         5.00         0        16.0  8.00  33418
 9  2000    3.00          1111 US000         5.00         0        16.0  9.00  30534
10  2000    4.00          1111 US000         5.00         0        16.0 10.0   30475
# ... with 54,074 more rows
```


## Backup: what if the BLS has changed my data!

I included an option to specify a snapshot of the data take through the url wayback machine: thanks To [Gabriel Chodorow-Reich](http://scholar.harvard.edu/chodorow-reich) for suggesting the idea.<sup>[1](#fn1)</sup>
Specify a suffix found from the internet wayback machine on the BLS website and add it as an argument in the download function:
```{r, results='hide', warning = F, error = F, message = F, eval = F}
# Enter the address directly from the html address
wayback_suffix <- "https://web.archive.org/web/20141101135821"  # suffix for November 1st, 2014
# Or enter the exact date at which date it has been crawled:
wayback_suffix <- 20141101135821  

# execute adding the wayback option
dt_naics <- get_files_cut(
  data_cut = 28,
  industry = "naics",
  year_start = 1990,  year_end = 2015,       # try not to ask for data from the future        
  url_wayback = wayback_suffix,
  path_data = "~/Downloads/", write = F)
dt_naics
```



# Documentation 

## <a name="tablecut">Cuts in QCEW</a>


For example the table of cuts looks as the following for **naics** based industries is as follows:

```{r, eval = F}
#------------------------------------------------------------------------------------------------------------------------------
#|                           | Geographic/Size   | National | National | CSA | MSA | Statewide | Statewide | County | MicroSA |
#|                           |  LevelCode Format |          |  by size |     |     |           |  by size  |        |         |
#|---------------------------|-------------------|----------|--- ------|-----|-----|-----------|-----------|--------|---------|
#| Industry Ownership Level  |                   | 1x       | 2x       | 3x  | 4x  | 5x        | 6x        | 7x     | 8x      |
#|                           |                   |          |          |     |     |           |           |        |         |
#| Total, All Ownerships     | x0                | 10       | -        | 30  | 40  | 50        | -         | 70     | 80      |
#| Total, by Ownership       | x1                | 11       | 21       | -   | 41  | 51        | 61        | 71     | -       |
#| Domain, by Ownership      | x2                | 12       | 22       | -   | 42  | 52        | 62        | 72     | -       |
#| SuperSector, by Ownership | x3                | 13       | 23       | -   | 43  | 53        | 63        | 73     | -       |
#| Sector, by Ownership      | x4                | 14       | 24       | -   | 44  | 54        | 64        | 74     | -       |
#| 3-digit, by Ownership     | x5                | 15       | 25       | -   | 45  | 55        | -         | 75     | -       |
#| 4-digit, by Ownership     | x6                | 16       | 26       | -   | 46  | 56        | -         | 76     | -       |
#| 5-digit, by Ownership     | x7                | 17       | 27       | -   | 47  | 57        | -         | 77     | -       |
#| 6-digit, by Ownership     | x8                | 18       | 28       | -   | 48  | 58        | -         | 78     | -       |
#------------------------------------------------------------------------------------------------------------------------------
```


## <a name = "generaldoc">General documentation</a>

The table of contents to download directly datasets

The data file layout for a general view of what is available:
  
  - For [downloading](https://www.bls.gov/cew/datatoc.htm) directly the individual files
  - For [table of contents](https://www.bls.gov/cew/datatoc.htm)
  - For [quarterly naics](http://www.bls.gov/cew/doc/layouts/csv_quarterly_layout.htm) and [annual naics](https://data.bls.gov/cew/doc/layouts/csv_annual_layout.htm)
  - For [quarterly sic](http://www.bls.gov/cew/doc/layouts/sic_csv_quarterly_layout.htm) and [annual sic](https://data.bls.gov/cew/doc/layouts/sic_csv_annual_layout.htm)
  



### Data codes

#### Aggregation levels

- Aggregation levels and files that contain them are defined by the BLS; we reproduce the table in the package for merging or easier access:
    + **naics** [table](../data_raw/naics_agglevel.csv); original [BLS webpage](http://www.bls.gov/cew/doc/titles/agglevel/agglevel_titles.htm)
    + **sic** [table](../data_raw/sic_agglevel.csv); original [BLS webpage](http://www.bls.gov/cew/doc/titles/agglevel/sic_agglevel_titles.htm)
    + **Supersectors** and **Sector**: head over to the [BLS webpage](https://www.bls.gov/cew/supersector.htm)
- Note that there is no disaggregated data for size X industry before 1990 in the case of **naics**. The layout of files on the [download page](http://www.bls.gov/cew/datatoc.htm) is somewhat misleading.

#### Size Classes
  - For [naics](http://www.bls.gov/cew/doc/titles/size/size_titles.htm)
  - For [sic](http://www.bls.gov/cew/doc/titles/size/sic_size_titles.htm)


#### Industries

Industry titles are standard in that case

  - **naics** [table](../data_raw/naics_industry_titles.csv); original [BLS webage](http://www.bls.gov/cew/doc/titles/industry/industry_titles.htm)
  - **sic** [table](../data_raw/sic_industry_titles.csv); original [BLS webpage](http://www.bls.gov/cew/doc/titles/industry/sic_industry_titles.htm)


#### Ownership  
Ownership codes go from 0 to 6:

  - 0	represents the aggregate or `Total Covered`
  - 5 represents the `Private` sector
  - 1 to 4 represent different level of government: 4	for `International Government`; 3	for `Local Government`; 2	for `State Government` and 1	for `Federal Government`

Online docs for [naics](http://www.bls.gov/cew/doc/titles/ownership/ownership_titles.htm) and [sic](http://www.bls.gov/cew/doc/titles/ownership/sic_ownership_titles.htm)


#### Area codes and titles (FIPS)
Structures somewhat like industry codes. 5 characters. `US000` is aggregated over the total US.
Then for example `XXYYY` can be split in two parts:

  - `XX` represents the state as in Census codes (alphabetical orders): 01 is Alabama and 02 is Alaska.
  - `YYY` represents the county within the given state
  - There are exceptions to `YYY` matching to counties:
    + If `YYY` is `000` then this represents data aggregated at the state level: 01000,
    + `YYY` is `996` then it represents	"Overseas Locations"
    + `YYY` is `997` then it represents "Multicounty, Not Statewide"
    + `YYY` is `998` then it represents "Out-of-State"
    + `YYY` is `999` then it represents "Unknown Or Undefined"
  - If the first character is `C` then it represents subdivisions at the MSA level

BLS documentation is available at the following links:

  - For [naics](http://www.bls.gov/cew/doc/titles/area/area_titles.htm)
  - For [sic](http://www.bls.gov/cew/doc/titles/area/sic_area_titles.htm)



---------------------------
<a name="fn1">1</a>: The `stata` version of this code is on Gabe's website [here](http://scholar.harvard.edu/chodorow-reich/data-programs)

---------------------------
(c) Erik Loualiche
