---
title: "Load from the BEA Survey of Current Business"
author: "Erik Loualiche"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
#output:
  #md_document:
  #  variant: markdown_github
vignette: >
  %\VignetteIndexEntry{Load from the BEA Survey of Current Business}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The dataset is small so it is available directly with the package (no download required). It includes three variables:

1. `nbf`: Index of net business formation (1967=100)
2. `nbi`: Number of new business incorporations (number)
3. `cbf`: Current liabilities of business failures, NSA (mil. $)


Load it with:
```{r, warning = F, error = F, message = F}
entrydatar::scb
```

There is also a function like the other in this package to access an interval:
```{r, warning = F, error = F, message = F}
entrydatar::get_nbf(1980, 1998)
```

Note that the data only extends to 1994 (included).

Quick summary stats of the dataset"
```{r, warning = F, error = F, message = F}
skimr::skim(entrydatar::get_nbf(1980, 1998)[, .(nbi, nbf, cbf)])
```






---------------------------



From the [original file](https://www.bea.gov/scb/pdf/NATIONAL/BUSCYCLE/1996/0296cpgs.pdf), we have the sources of the information gathered here:

  + Mr. Neil DiBernardo (new business incorporations and business failures), *The Dun & Bradstreet Corporation, Economic Analysis Department*
  (Other component data are not available to the public.)
   
  
  

---------------------------
(c) Erik Loualiche
